\chapter{Assignment 4: Hypothesis Test}

\section*{$\xi$ values}

\begin{itemize}
    \item $\xi_{11} = 912$
    \item $\xi_{12} = 36.6$
    \item $\xi_{13} = 2$
    \item $\xi_{14} = [879, 842, 954, 842, 885, 918, 989, 768, 867, 1022]$
\end{itemize}


\section{Problem statement}

A statistical hypothesis testing is needed to determine if a new production system for hammers yields higher weights than the current system. The weights of hammers produced in the factory are normally distributed with a mean of 912 grams and a standard deviation of 36.6 grams.
The analysis steps include setting up the hypotheses, performing the test, and making a decision based on the results.

\subsection*{Proposed Model for the Hammer Weights}

The weights of hammers produced in the factory can be modeled using a normal distribution based on the observed long-term data:
\begin{equation}
W \sim \mathcal{N}(\mu, \sigma^2)
\end{equation}
where:
\begin{itemize}
    \item $\mu = \xi_{11}$ is the mean weight of the hammers.
    \item $\sigma = \xi_{12}$ is the standard deviation of the weights.
\end{itemize}

\subsubsection*{Assumptions}
The following assumptions are made for this model to hold:
\begin{itemize}
    \item The weights of the hammers are independent .
    \item mean and standard deviation are not varying in time.
    \item The underlying distribution of weights is approximately normal
\end{itemize}

\subsubsection*{Model Parameters}
The parameters of the model are:
\begin{itemize}
    \item $\mu$: Mean of the distribution.
    \item $\sigma^2$: Variance of the distribution (or $\sigma$: standard deviation).
\end{itemize}

\section{Hypothesis Testing}

\subsection*{chosen statistical test and Decision Rule}
I perform a one-sample one-tailed $ t $-test


The test statistic is given by:
\begin{equation}
t = \frac{\bar{x} - \mu_0}{s / \sqrt{n}}
\end{equation}
where:
\begin{itemize}
    \item $\bar{x}$ is the sample mean,
    \item $s$ is the sample standard deviation,
    \item $n$ is the sample size,
    \item $\mu_{0} = 912$ is the null hypothesis mean.
\end{itemize}

The decision rule is to reject $H_{0}$ if $t>t_{c}$, where $ t_{c} $ is determined from the $ t $-distribution table at a chosen significance level $\alpha$ with $ n-1 = 9 $ degrees of freedom.

To determine if the new production system yields higher weights, we set up the following hypotheses:
\begin{itemize}
\item $H_{0}$ and $\mu = 912$ : the mean weight remains unchanged
\item $H_{a}$ and $\mu > 912$ :  the mean weight is higher under the new system
\end{itemize}


\subsection*{Error Probabilities}
\begin{itemize}
    \item Type I error ($\alpha$): Rejecting $H_0$ when $H_0$ is true. Choose $\alpha = 0.05$.
    \item Type II error ($\beta$): Failing to reject $H_0$ when $H_a$ is true. Can be estimated if the true mean under $H_a$ is known.
\end{itemize}

\subsection*{test computation}
The sample weights are:
\begin{equation}
[879, 842, 954, 842, 885, 918, 989, 768, 867, 1022]
\end{equation}

Calculate the sample mean:
\begin{equation}
\bar{x} = \frac{\sum x_i}{n} = \frac{879 + 842 + \cdots + 1022}{10} = 896.6
\end{equation}

Calculate the sample standard deviation:
\begin{equation}
s = \sqrt{\frac{1}{n-1} \sum_{i=1}^n (x_i - \bar{x})^2} \approx 71.5
\end{equation}

Compute the test statistic:
\begin{equation}
t = \frac{\bar{x} - \mu_0}{s / \sqrt{n}} = \frac{896.6 - 912}{71.5 / \sqrt{10}} \approx -0.69
\end{equation}

\subsection*{Decision and Conclusion}
For $\alpha = 0.05$ and $df = 9$, $t_{c} \approx 1.8$.
Since $t = -0.69 < t_{c}$, I fail to reject $H_{0}$.

\textbf{Conclusion}: There is insufficient evidence to suggest that the new system produces hammers with higher weights.
